Machine Learning Project: Predicting how well individual performs  barbell lifts
==================================================================

#Introduction
Devices such as Jawbone Up, Nike FuelBand, and Fitbit provide a large amount of data about personal activity.  This analysis uses data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants, who were asked to perform barbell lifts correctly and incorrectly in 5 different ways.  

#Load and Explore Data
```{r,warning=FALSE,eval=FALSE}
library(caret)
dataset <- read.csv("train.csv")
inTrain <- createDataPartition(dataset$classe,p=0.7,list=FALSE)
training <- dataset[inTrain,]
testing <- dataset[-inTrain,]
str(training)
```

#Prepreocess Data
There are 160 variables.  After careful examination of the variable, some of the variables are statistical properties such as
kurtosis, min, max, etc.  I will remove thses variables since they are not predictive variables

```{r,warning=FALSE,eval=FALSE}
#identify columns with statistical properties
stats_id <- c("kurtosis","skewness","max","min","amplitude","avg","stddev","var")
col_remove <-c()
for(i in stats_id){
  k <- grep(i, names(training), value = F)
  col_remove <- c(col_remove, as.integer(k))
}

#The first 7 columns are housekeeping items and also need to be removed
col_remove <- c(c(1:7),col_remove)

#Remove unwanted columns
training <- training[,-col_remove]
testing <- testing[,-col_remove]
dim(training)
dim(testing)
```

#Model Fitting 

After data preprocess, 56 covariates are left in the dataset, but I suspect many of them are correlated. 
Therefore, I use PCA before fitting a model using random forest. 

```{r,warning=FALSE,eval=FALSE}
modFit <- train(classe~., method="rf",preProcess="pca",data=training)
```

#Model Evaluation

The in sample error is 0. However, the out of sample error should be greater than 0.
```{r,warning=FALSE,eval=FALSE}
confusionMatrix(training$classe,predict(modFit))
```

```{r,warning=FALSE,eval=FALSE}
##Confusion Matrix and Statistics

          Reference
Prediction    A    B    C    D    E
         A 3906    0    0    0    0
         B    0 2658    0    0    0
         C    0    0 2396    0    0
         D    0    0    0 2252    0
         E    0    0    0    0 2525

Overall Statistics
                                     
               Accuracy : 1          
                 95% CI : (0.9997, 1)
    No Information Rate : 0.2843     
    P-Value [Acc > NIR] : < 2.2e-16  
                                     
                  Kappa : 1          
 Mcnemar's Test P-Value : NA         

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            1.0000   1.0000   1.0000   1.0000   1.0000
Specificity            1.0000   1.0000   1.0000   1.0000   1.0000
Pos Pred Value         1.0000   1.0000   1.0000   1.0000   1.0000
Neg Pred Value         1.0000   1.0000   1.0000   1.0000   1.0000
Prevalence             0.2843   0.1935   0.1744   0.1639   0.1838
Detection Rate         0.2843   0.1935   0.1744   0.1639   0.1838
Detection Prevalence   0.2843   0.1935   0.1744   0.1639   0.1838
Balanced Accuracy      1.0000   1.0000   1.0000   1.0000   1.0000
```

As expected, the out of sample error is 0.028, which is greater than 0.
```{r,warning=FALSE,eval=FALSE}
confusionMatrix(testing$classe,predict(modFit,testing))
```

```{r,warning=FALSE,eval=FALSE}
##Confusion Matrix and Statistics

          Reference
Prediction    A    B    C    D    E
         A 1661    3    2    7    1
         B   26 1093   19    0    1
         C    2   11  998   15    0
         D    1    0   46  914    3
         E    2    5   13    8 1054

Overall Statistics
                                         
               Accuracy : 0.972          
                 95% CI : (0.9674, 0.976)
    No Information Rate : 0.2875         
    P-Value [Acc > NIR] : < 2.2e-16      
                                         
                  Kappa : 0.9645         
 Mcnemar's Test P-Value : NA             

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            0.9817   0.9829   0.9258   0.9682   0.9953
Specificity            0.9969   0.9904   0.9942   0.9899   0.9942
Pos Pred Value         0.9922   0.9596   0.9727   0.9481   0.9741
Neg Pred Value         0.9926   0.9960   0.9835   0.9939   0.9990
Prevalence             0.2875   0.1890   0.1832   0.1604   0.1799
Detection Rate         0.2822   0.1857   0.1696   0.1553   0.1791
Detection Prevalence   0.2845   0.1935   0.1743   0.1638   0.1839
Balanced Accuracy      0.9893   0.9866   0.9600   0.9791   0.9947
```


#Prediction for the 20 test cases

```{r,warning=FALSE,eval=FALSE}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile="test.csv",method="curl")
test_pred <- read.csv("test.csv")
test_pred <- test_pred[,-col_remove]
prediction <- predict(modFit,test_pred)
```

Prediction output: B A B A A E D B A A B C B A E E A B B B